Updated Constraints on Self-Interacting Dark Matter from Supernova 1987A

### 2017-01-10 draft.tex
Made changes all the way through to the results section.

\newcommand{\as}{\alpha_s}
\newcommand{\qa}{q_{A'}}

@@

@@

SIDM models in which large self-interaction cross sections are mediated by sufficiently light 
bosons ($M \lesssim 100$~GeV) can be constrained astrophysically using supernovae, particularly 
SN1987A. Light gauge bosons will be produced within the hot supernovae core and 
radiated from the supernova. This non-standard energy loss mechanism can result in an energy loss rate 
from the supernova core that is inconsistent with observations of SN1987A, analogous to the 
classic constraint on axions \cite{raffelt96_book}. This mechanism was already exploited by 
\cite{dent_etal12} (and \cite{rrapaj_reddy16}, with a slight modification) (\arz{Cameron, add others of which you may be aware.}) 
to constraint dark electromagnetism models of SIDM in which the self-interaction arises from a Yukawa potential 
and the vector boson is kinetically mixed with the standard model photon.

As we mentioned above, we find SN1987A constraints on dark photon SIDM models that are considerably different 
from those of \cite{dent_etal12}. If correct, our result has important consequences. 
The Snowmass white paper by Kaplinghat, Tulin, and Yu \cite{kaplinghat_etal13_whitepaper} nicely 
summarizes a variety of constraints on dark electromagnetism models of SIDM. 
One point that is clear from Ref.~\cite{kaplinghat_etal13_whitepaper} is that there exists 
only a narrow range of viable photon mixing parameters that can lead to production of SIDM in the early universe and 
interesting effects on cosmological structure formation, while simultaneously evading all constraints, 
including the constraint of Ref.~\cite{dent_etal12}. Our results suggest that the SN1987A constraints from 
Ref.~\cite{dent_etal12} are too restrictive. The constraints that we quote reveal a region of unrestricted mixing 
parameters that ranges over two orders of magnitude. Interestingly, the parameter space opened by our updated 
constraints corresponds to SIDM models that may have a significant impact on cosmological structure formation, 
particularly the structures of dark matter halos and galaxies.

@@

\arz{Need to add citations to many of the Feng et al. papers here.}
We consider constraints on SIDM specifically within the context of dark electromagnetism models. 
Dark electromagnetism models are models in which a hidden, dark, sector contains a broken U(1)$'$ 
symmetry and the U(1)$'$ gauge boson is kinetically mixed with the standard model photon. For the 
purposes of this study, this is important because it demands that the Lagrangian contains terms such as 
%
\begin{equation}
\mathcal{L}_{\mathrm{int}} = g_{\chi} \bar{\chi} \, \widetilde{\slashed{A'}} \chi + q \bar{f}\, \widetilde{\slashed{A}} f, 
\end{equation}
%
where $\chi$ is the dark matter, $g_{\chi}$ is the dark coupling, $\widetilde{A'}$ is the dark gauge boson, 
$f$ is a standard model fermion of charge $q$, and $\widetilde{A}$ is the standard model gauge boson. 
The kinetic mixing, through a term 
$\frac{1}{2}\frac{\varepsilon}{\sqrt{1+\varepsilon^2}}\, \widetilde{F}_{\mu \nu}\widetilde{F}'^{\mu \nu}$ 
in the Lagrangian causes the $\widetilde{A}$ to be an admixture of the massless photon $A$, 
and the dark photon $A'$, 
of mass $m_{\mathrm{A'}} = m_{\mathrm{\widetilde{A}'}} \sqrt{1 + \varepsilon^2} \simeq m_\mathrm{{\widetilde{A}'}}$ because 
the viable parameter range has $\varepsilon \ll 1$. The dark matter particles are thereby coupled 
to the standard model fermions with a coupling constant $\varepsilon q$, 
where $\varepsilon$ is the kinetic mixing parameter. The first term in this 
interaction Lagrangian gives rise to the dark matter self-interactions.

Dark gauge bosons are produced in astrophysical environments such as supernova 
cores primarily via brehmsstrahlung off of standard model particles. This brehmsstrahlung 
occurs through the $\varepsilon q$ coupling to charged standard model particles, 
in this  particular case the proton and pion. The rate of brehmsstrahlung 
depends upon both $\varepsilon$ and the mass of the $A'$. Consequently, 
supernova cooling can constrain the mixing $\varepsilon$ as a function of 
$m_{A'}$ for such models. Delineating such a constraint is the 
primary aim of this paper.

@@

We aim to estimate the rate of energy loss from the core of a supernova from $A'$ brehmsstrahlung 
during nucleon-nucleon interactions. The calculation is analogous to the well-known estimate of axion 
emission from supernova cores described in Ref.~\cite{raffelt96_book} and references therein, but is 
more complicated because the mass of the $A'$, unlike the mass of the axion, is not necessarily 
negligible. This section describes the calculation of the rate of energy loss from a supernova core 
from $A'$ brehmsstrahlung.

The brehmsstrahlung process is not the only process with which we must be concerned. Clearly, the rate of 
brehmsstrahlung will increase with $\varepsilon$; however, $\varepsilon$ can become sufficiently large 
that the radiated gauge bosons do not escape the supernova. This happens if the $A'$ either decay to 
standard model particles or interact with standard model particles prior to exiting the supernova core. 
In either case, the energy is not lost and the $A'$ does not provide a cooling channel for the supernova. 
Consequently, for a given $m_{A'}$, there is a maximum $\varepsilon$ that can be constrained in this 
manner. We estimate $A'$ decay and scattering probabilities, and the upper limits on the $\varepsilon$ 
constraints in this section as well.

\subsection{Brehmsstrahlung of $A'$ Bosons}

There are two processes to consider in order to estimate the rate of energy loss via $A'$ brehmsstrahlung. The first is  
proton-proton (pp) scattering with the bremsstrahlung of the dark photon off the proton; $p+p \rightarrow p+p+A$. The 
second is proton-neutron (pn) scattering with bremsstrahlung off of either the proton or the charged pion; 
$p+n \rightarrow p+n+A$. We estimate the rates for these processes using the one-pion exchange (OPE) 
approximation for nucleon interactions. In the pp case, there are eight tree-level diagrams, 
with the emission of the $A'$ from each of the external legs. One of these diagrams is shown 
in Fig.~\ref{fig:ppdiagram}; the remaining seven diagrams come from placing the radiated $A'$ 
on each of the other three protons and then, for each of these, interchanging the outgoing momenta. 
For the pn case, there are five diagrams, four of which are analogous to the pp diagram shown in 
Fig.~\ref{fig:ppdiagram}. The fifth diagram, shown in Fig.~\ref{fig:npdiagram}, corresponds to 
emission of the $A'$ from the exchanged, charged pion.

%-------------- PP Feynman Diagram
\begin{figure}
\includegraphics[width=8cm]{ppdiagram.pdf}
\caption{One of the eight Feynman diagrams for the pp process. Three of the other diagrams are obtained by 
placing the $A'$ on each of the protons in turn. The remaining four diagrams come from swapping the outgoing 
momenta.}
\label{fig:ppdiagram}
\end{figure}

Evaluating these diagrams is tedious, but very straightforward. The calculation differs 
from the well-known axion bremsstrahlung calculation, because the mass of the $A'$ boson is not 
necessarily negligible in the kinematic region of interest for supernova explosions.
The correct kinematical relations are 
%
\bea 
p_1 \cdot p_2 &=& M_N^2 - \frac{l^2}{2} - \frac{k^2}{2} + p_2 \cdot \qa,\\
p_1 \cdot p_3 &=& M_N^2 + k \cdot l - \frac{k^2}{2} + p_3 \cdot \qa,\\  
p_1 \cdot p_4 &=& k \cdot l + M_N^2 - \frac{l^2}{2} + p_4 \cdot \qa, \\
p_2 \cdot p_3 &=& M_N^2 - \frac{l^2}{2}, \\ 
p_2 \cdot p_4 &=& M_N^2 - \frac{k^2}{2},\quad \mathrm{and}\\
p_3 \cdot p_4 &=& k \cdot l + M_N^2 - \frac{l^2 + k^2}{2},
\eea
%
where $p_1$ and $p_2$ are the four-momenta of the incoming nucleons, $p_3$ and $p_4$ are the momenta 
of the outgoing nucleons, $\qa$ is the $A'$ momentum, $k=p_2 - p_3$, $l=p_2 - p_4$, and $M_{\mathrm{N}}$ 
is the nucleon mass. These kinematical relations correct the relations in Ref.~\cite{dent_etal12}.

@@

The pn process contain four diagrams analogous to the the diagrams for the pp process (there are only four, of course, 
because the neutrons cannot radiate the $A'$). The new diagram that is relevant in the pn process is shown in 
Fig.~\ref{fig:npdiagram} and yields a contribution of 
%
\begin{widetext}
\begin{equation}
M'_5 = \frac{4 M_N}{ m_\pi} \frac{f_{pn}^2 e\, \varepsilon}{l^2-m_\pi^2}  \frac{1}{(l-\qa)^2 - m_\pi^2} 
\bar{u}(p_3) \gamma_5 u(p_1) \bar{u}(p_4) \gamma_5 u(p_2) (\qa - 2l)\cdot \varepsilon.
\end{equation}
\end{widetext}
%
The pn processes likewise yield a squared amplitude, 
$\vert \mathcal{M}_{\mathrm{pn}} \vert^2$ that is unwieldy, so we do not give the 
the full expression here.

@@

The first and simplest bound that may be obtained arises from assuming that all $A'$ particles produced in the supernova core 
leave the supernova, carrying their energies with them. The constraint can be derived simply by 
requiring that the energy loss through this cooling channel be less than the cooling from neutrino emission; 
any greater, and it would have an observable effect on supernova cooling. This calculation yields values of 
$\varepsilon$ above which cooling through $A'$ production is too rapid to be consistent with SN1987A. 
We will consider modifications to this bound from $A'$ trapping and decay in subsequent subsections.

The quantity of interest is the rate of energy emission through dark gauge bosons. 
From the spin-summed, squared amplitudes described in the previous subsection, 
the energy emission rate is obtained by integrating over the phase space, 
and adding a factor of the energy of the emitted particle. To be specific, the 
energy emission rate per unit volume is 
%
\begin{widetext}
\beq
\label{eq:rate1}
Q_i = (2\pi)^4 \int\, E_{\mathrm{A}'} \, \sum_{s_1,s_2}\, \vert \mathcal{M}_i \vert^2 f(p_1) f(p_2)\delta^{(4)}(p_1+p_2-p_3-p_4-\qa)\, \dd \Pi,
\eeq
\end{widetext}
%
where 
%
\begin{equation}
d\Pi = \frac{\dd^3 \vec{q}_{\mathrm{A}'}}{(2\pi)^3 2E_{\mathrm{A}'}}\, \prod_{i=1}^{4}\, \frac{\dd^3 \vec{p}_i}{(2\pi^3) 2E_1}
\end{equation} 
is the Lorentz-invariant phase space interval, 
$ E_{\mathrm{A}'}$ is the energy of the emitted $A'$ boson, 
$f(p)$ are the phase-space densities of the incoming nucleons, and 
the index $i$ refers to either the pp or pn processes. 
The nucleons in the core are comfortably non-degenerate and non-relativistic, 
so the Pauli blocking factor is omitted from Eq.~(\ref{eq:rate1}) and we take 
all nucleons to have a Maxwell-Boltzmann phase-space distribution distribution, 
\beq
f(p) =  \frac{n_b}{2} \left( \frac{2 \pi}{M_N T} \right)^{3/2} \exp \left(-\frac{p^2}{2 M_N T} \right).
\eeq
%
We choose a baryon number density of $n_{\mathrm{b}} \approx 1.8 \times 10^{38}\, \mathrm{cm}^{-3}$ and a 
core supernova temperature of $T = 30\, \mathrm{MeV}$, both of which are typical choices and thought to 
be representative of supernova cores. \arz{Need to put a citation here for the properties of supernova cores. 
I can probably find something in the literature, but please put something here if you have a citation.}

@@

To obtain the dark gauge boson luminosity from $Q_{\mathrm{pp}}$ and $Q_{\mathrm{pn}}$, 
we assumed that $A'$ production takes place in a stellar core of radius $\sim 1$~km, so that 
the total luminosity of $A'$ is 
%
\begin{equation}
L_{\mathrm{A}'} = V (Q_{\mathrm{pp}} + Q_{\mathrm{pn}})
\end{equation}
%
where $V$ is the volume of the sphere. Clearly, $L_{\mathrm{A}'}$ is proportional to $\varepsilon^2$. 
Writing $L_{\mathrm{A}'} = \varepsilon^2 I_{\mathrm{A}'}(m_{\mathrm{A}'},T)$, as in Ref.~\cite{dent_etal12}, 
and $L_{\nu}$ as the neutrino luminosity, we can write the constraint as 
%
\begin{equation}
\label{eq:bound1}
\varepsilon \lesssim \sqrt{\frac{L_{\nu}}{I_{\mathrm{A}'}(m_{\mathrm{A}'}, T)}}.
\end{equation}
%
Following SOMEBODY \arz{We need a citation here.}, we derive a constraint taking 
$L_{\nu} \approx 10^{53}\, \mathrm{erg/s} \approx 4 \times 10^{37}\, \mathrm{MeV}^2$.

The constraint derived in this manner from Eq.~(\ref{eq:bound1}) sets the lower limit on the 
exclusion band shown in Figure~\ref{fig:constraint}. We will discuss Fig.~\ref{fig:constraint} 
in more detail below. If all of the $A'$ produced in the core leave the supernova freely, all 
values of $\varepsilon$ higher than those given by Eq.~(\ref{eq:bound1}) would be ruled out. 
However, as we have already mentioned, $\varepsilon$ can become sufficiently large that only a 
negligible amount of energy actually exits the supernova core. For large values or $\varepsilon$, 
this can occur because of either $A'$ decays or $A'$ scattering. These additional considerations place 
an upper limit on the values of $\varepsilon$ for which this constraint applies, 
and we discuss these effects in the next two subsections.

@@

The constraint from the streaming limit above provides an upper bound on the allowed coupling 
from the production on $A'$ bosons. In order to function as an effective cooling channel, the 
radiated dark gauge bosons must escape the supernova in sufficient numbers. 
As the coupling increases, so do processes that prevent this escape, 
and so the excluded region has an upper bound, above which the luminosity again drops below the 
neutrino luminosity. The first such limit may be found by considering decay of the dark bosons 
into Standard Model particles. Standard model particles will scatter and thermalize 
on a timescale much shorter than the timescale for the evolution of the core, so decays 
contribute nothing to the cooling.

\arz{I removed reference to $N_{\rm eff}$ because for us $N_{\rm eff}=1$ (doesn't it? the 
only kinematically-allowed decay is to electron-positron pairs, correct?). I also rephrased a 
little bit. Please review this. 
The dark boson has a typical lifetime of
\begin{equation}
\tau_{\mathrm{A}'} = \frac{3}{\varepsilon^2 \alpha m_{\mathrm{A}'}},
\end{equation}
where $\alpha$ is the fine structure constant. Therefore, in the relativistic limit, 
the typical travel distance to decay is given by 
%
\beq
l = \gamma \tau_{\mathrm{A}'} = \frac{3 E_{A}}{N_{eff} m_A^2 \varepsilon^2},
\eeq
and so the fraction escaping the supernova before decaying may be estimated as (e.g., \Ref.~\cite{bjorken_etal09})
\beq
e^{r_{decay}/l} = e^{r_{decay} N_{eff} m_A^2 \varepsilon^2/(3 E_A)}.
\eeq
%
To account for the suppression of gauge boson luminosity due to decays we simply multiply the phase space 
integrand in Eq.~(\ref{eq:rate1}) by the exponential suppression factor, after which the calculation proceeds 
as in the previous subsection. The limit is derived in the same way, 
with the additional complication that $I_{\mathrm{A}'}$ is now a function of $\varepsilon $, in addition to $m_{\mathrm{A}'}$ and $T$. 
The constraint Eq.~(\ref{eq:bound1}), therefore, becomes a transcendental equation that must be solved numerically.

The luminosity in $A'$ will be an increasing function of $\varepsilon$ until decays suppress the gauge boson luminosity, 
at which point $L_{\mathrm{A}'}$ becomes a rapidly decreasing function of $\varepsilon$. Therefore, the excluded values of 
$\varepsilon$ at a given mass will generally have a lower bound set by the calculations of the previous subsection, and an 
upper bound set by decays.

@@

Let's put some text here.

>>>>

\newcommand{\as}{\alpha_s}
\newcommand{\qa}{q_{\mathrm{A}'}}

@@

\newcommand{\Aprime}{\mathrm{A}'}
\newcommand{\rdp}{r_{\mathrm{dp}}}

@@

SIDM models in which large self-interaction cross sections are mediated by sufficiently light 
bosons ($M \lesssim 100$~GeV) can be constrained astrophysically using supernovae, particularly 
SN1987A. Light gauge bosons will be produced within the hot supernovae core and 
radiated from the supernova. This non-standard energy loss mechanism can result in an energy loss rate 
from the supernova core that is inconsistent with observations of SN1987A, analogous to the 
classic constraint on axions \cite{turner88,raffelt96_book}. This mechanism was already exploited by 
\cite{dent_etal12} (and \cite{rrapaj_reddy16}, with a slight modification) (\arz{Cameron, add others of which you may be aware.}) 
to constraint dark electromagnetism models of SIDM in which the self-interaction arises from a Yukawa potential 
and the vector boson is kinetically mixed with the standard model photon.

As we mentioned above, we find SN1987A constraints on dark photon SIDM models that are considerably different 
from those of \cite{dent_etal12}. If correct, our result has important consequences. 
The Snowmass white paper by Kaplinghat, Tulin, and Yu \cite{kaplinghat_etal13_whitepaper} nicely 
summarizes a variety of constraints on dark electromagnetism models of SIDM (many details can be 
found in the the very informative work of Ref.~\cite{bjorken_etal09}). 
One point that is clear from Ref.~\cite{kaplinghat_etal13_whitepaper} is that there exists 
only a narrow range of viable photon mixing parameters that can lead to production of SIDM in the early universe and 
interesting effects on cosmological structure formation, while simultaneously evading all constraints, 
including the constraint of Ref.~\cite{dent_etal12}. Our results suggest that the SN1987A constraints from 
Ref.~\cite{dent_etal12} are too restrictive. The constraints that we quote reveal a region of unrestricted mixing 
parameters that ranges over two orders of magnitude. Interestingly, the parameter space opened by our updated 
constraints corresponds to SIDM models that may have a significant impact on cosmological structure formation, 
particularly the structures of dark matter halos and galaxies.

@@

\arz{Need to add citations to many of the Feng et al. papers here.}
We consider constraints on SIDM specifically within the context of dark electromagnetism models. 
Dark electromagnetism models are models in which a hidden, dark, sector contains a broken U(1)$'$ 
symmetry and the U(1)$'$ gauge boson is kinetically mixed with the standard model photon. For the 
purposes of this study, this is important because it demands that the Lagrangian contains terms such as 
%
\begin{equation}
\mathcal{L}_{\mathrm{int}} = g_{\chi} \bar{\chi} \, \widetilde{\slashed{\Aprime}} \chi + q \bar{f}\, \widetilde{\slashed{A}} f, 
\end{equation}
%
where $\chi$ is the dark matter, $g_{\chi}$ is the dark coupling, $\widetilde{\Aprime}$ is the dark gauge boson, 
$f$ is a standard model fermion of charge $q$, and $\widetilde{A}$ is the standard model gauge boson. 
The kinetic mixing, through a term 
$\frac{1}{2}\frac{\varepsilon}{\sqrt{1+\varepsilon^2}}\, \widetilde{F}_{\mu \nu}\widetilde{F}'^{\mu \nu}$ 
in the Lagrangian causes the $\widetilde{A}$ to be an admixture of the massless photon $A$, 
and the dark photon $\Aprime$, 
of mass $m_{\mathrm{\Aprime}} = m_{\mathrm{\widetilde{A}'}} \sqrt{1 + \varepsilon^2} \simeq m_\mathrm{{\widetilde{A}'}}$ because 
the viable parameter range has $\varepsilon \ll 1$. The dark matter particles are thereby coupled 
to the standard model fermions with a coupling constant $\varepsilon q$, 
where $\varepsilon$ is the kinetic mixing parameter. The first term in this 
interaction Lagrangian gives rise to the dark matter self-interactions.

Dark gauge bosons are produced in astrophysical environments such as supernova 
cores primarily via brehmsstrahlung off of standard model particles. This brehmsstrahlung 
occurs through the $\varepsilon q$ coupling to charged standard model particles, 
in this  particular case the proton and pion. The rate of brehmsstrahlung 
depends upon both $\varepsilon$ and the mass of the $\Aprime$. Consequently, 
supernova cooling can constrain the mixing $\varepsilon$ as a function of 
$m_{\Aprime}$ for such models. Delineating such a constraint is the 
primary aim of this paper.

@@

We aim to estimate the rate of energy loss from the core of a supernova from $\Aprime$ brehmsstrahlung 
during nucleon-nucleon interactions. The calculation is analogous to the well-known estimate of axion 
emission from supernova cores described in Ref.~\cite{raffelt96_book} and references therein, but is 
more complicated because the mass of the $\Aprime$, unlike the mass of the axion, is not necessarily 
negligible. This section describes the calculation of the rate of energy loss from a supernova core 
from $\Aprime$ brehmsstrahlung.

The brehmsstrahlung process is not the only process with which we must be concerned. Clearly, the rate of 
brehmsstrahlung will increase with $\varepsilon$; however, $\varepsilon$ can become sufficiently large 
that the radiated gauge bosons do not escape the supernova. This happens if the $\Aprime$ either decay to 
standard model particles or interact with standard model particles prior to exiting the supernova core. 
In either case, the energy is not lost and the $\Aprime$ does not provide a cooling channel for the supernova. 
Consequently, for a given $m_{\Aprime}$, there is a maximum $\varepsilon$ that can be constrained in this 
manner. We estimate $\Aprime$ decay and scattering probabilities, and the upper limits on the $\varepsilon$ 
constraints in this section as well.

\subsection{Brehmsstrahlung of $\Aprime$ Bosons}

There are two processes to consider in order to estimate the rate of energy loss via $\Aprime$ brehmsstrahlung. The first is  
proton-proton (pp) scattering with the bremsstrahlung of the dark photon off the proton; $p+p \rightarrow p+p+A$. The 
second is proton-neutron (pn) scattering with bremsstrahlung off of either the proton or the charged pion; 
$p+n \rightarrow p+n+A$. We estimate the rates for these processes using the one-pion exchange (OPE) 
approximation for nucleon interactions. In the pp case, there are eight tree-level diagrams, 
with the emission of the $\Aprime$ from each of the external legs. One of these diagrams is shown 
in Fig.~\ref{fig:ppdiagram}; the remaining seven diagrams come from placing the radiated $\Aprime$ 
on each of the other three protons and then, for each of these, interchanging the outgoing momenta. 
For the pn case, there are five diagrams, four of which are analogous to the pp diagram shown in 
Fig.~\ref{fig:ppdiagram}. The fifth diagram, shown in Fig.~\ref{fig:npdiagram}, corresponds to 
emission of the $\Aprime$ from the exchanged, charged pion.

%-------------- PP Feynman Diagram
\begin{figure}
\includegraphics[width=8cm]{ppdiagram.pdf}
\caption{One of the eight Feynman diagrams for the pp process. Three of the other diagrams are obtained by 
placing the $\Aprime$ on each of the protons in turn. The remaining four diagrams come from swapping the outgoing 
momenta.}
\label{fig:ppdiagram}
\end{figure}

Evaluating these diagrams is tedious, but very straightforward. The calculation differs 
from the well-known axion bremsstrahlung calculation, because the mass of the $\Aprime$ boson is not 
necessarily negligible in the kinematic region of interest for supernova explosions.
The correct kinematical relations are 
%
\bea 
p_1 \cdot p_2 &=& M_N^2 - \frac{l^2}{2} - \frac{k^2}{2} + p_2 \cdot \qa,\\
p_1 \cdot p_3 &=& M_N^2 + k \cdot l - \frac{k^2}{2} + p_3 \cdot \qa,\\  
p_1 \cdot p_4 &=& k \cdot l + M_N^2 - \frac{l^2}{2} + p_4 \cdot \qa, \\
p_2 \cdot p_3 &=& M_N^2 - \frac{l^2}{2}, \\ 
p_2 \cdot p_4 &=& M_N^2 - \frac{k^2}{2},\quad \mathrm{and}\\
p_3 \cdot p_4 &=& k \cdot l + M_N^2 - \frac{l^2 + k^2}{2},
\eea
%
where $p_1$ and $p_2$ are the four-momenta of the incoming nucleons, $p_3$ and $p_4$ are the momenta 
of the outgoing nucleons, $\qa$ is the $\Aprime$ momentum, $k=p_2 - p_3$, $l=p_2 - p_4$, and $M_{\mathrm{N}}$ 
is the nucleon mass. These kinematical relations correct the relations in Ref.~\cite{dent_etal12}.

@@

The pn process contain four diagrams analogous to the the diagrams for the pp process (there are only four, of course, 
because the neutrons cannot radiate the $\Aprime$). The new diagram that is relevant in the pn process is shown in 
Fig.~\ref{fig:npdiagram} and yields a contribution of 
%
\begin{widetext}
\begin{equation}
M'_5 = \frac{4 M_N}{ m_\pi} \frac{f_{pn}^2 e\, \varepsilon}{l^2-m_\pi^2}  \frac{1}{(l-\qa)^2 - m_\pi^2} 
\bar{u}(p_3) \gamma_5 u(p_1) \bar{u}(p_4) \gamma_5 u(p_2) (\qa - 2l)\cdot \varepsilon.
\end{equation}
\end{widetext}
%
The pn processes likewise yield a squared amplitude, 
$\vert \mathcal{M}_{\mathrm{pn}} \vert^2$ that is unwieldy, so we do not give the 
the full expression here.

@@

The first and simplest bound that may be obtained arises from assuming that all $\Aprime$ particles produced in the supernova core 
leave the supernova, carrying their energies with them. The constraint can be derived simply by 
requiring that the energy loss through this cooling channel be less than the cooling from neutrino emission; 
any greater, and it would have an observable effect on supernova cooling. This calculation yields values of 
$\varepsilon$ above which cooling through $\Aprime$ production is too rapid to be consistent with SN1987A. 
We will consider modifications to this bound from $\Aprime$ trapping and decay in subsequent subsections.

The quantity of interest is the rate of energy emission through dark gauge bosons. 
From the spin-summed, squared amplitudes described in the previous subsection, 
the energy emission rate is obtained by integrating over the phase space, 
and adding a factor of the energy of the emitted particle. To be specific, the 
energy emission rate per unit volume is 
%
\begin{widetext}
\beq
\label{eq:rate1}
Q_i = (2\pi)^4 \int\, E_{\mathrm{A}'} \, \sum_{s_1,s_2}\, \vert \mathcal{M}_i \vert^2 f(p_1) f(p_2)\delta^{(4)}(p_1+p_2-p_3-p_4-\qa)\, \dd \Pi,
\eeq
\end{widetext}
%
where 
%
\begin{equation}
d\Pi = \frac{\dd^3 \vec{q}_{\mathrm{A}'}}{(2\pi)^3 2E_{\mathrm{A}'}}\, \prod_{i=1}^{4}\, \frac{\dd^3 \vec{p}_i}{(2\pi^3) 2E_1}
\end{equation} 
is the Lorentz-invariant phase space interval, 
$ E_{\mathrm{A}'}$ is the energy of the emitted $\Aprime$ boson, 
$f(p)$ are the phase-space densities of the incoming nucleons, and 
the index $i$ refers to either the pp or pn processes. 
The nucleons in the core are comfortably non-degenerate and non-relativistic, 
so the Pauli blocking factor is omitted from Eq.~(\ref{eq:rate1}) and we take 
all nucleons to have a Maxwell-Boltzmann phase-space distribution distribution, 
\beq
f(p) =  \frac{n_b}{2} \left( \frac{2 \pi}{M_N T} \right)^{3/2} \exp \left(-\frac{p^2}{2 M_N T} \right).
\eeq
%
We choose a baryon number density of $n_{\mathrm{b}} \approx 1.8 \times 10^{38}\, \mathrm{cm}^{-3}$ and a 
core supernova temperature of $T = 30\, \mathrm{MeV}$, both of which are typical choices and thought to 
be representative of supernova cores. \arz{Need to put a citation here for the properties of supernova cores. 
I can probably find something in the literature, but please put something here if you have a citation.}

@@

To obtain the dark gauge boson luminosity from $Q_{\mathrm{pp}}$ and $Q_{\mathrm{pn}}$, 
we assumed that $\Aprime$ production takes place in a stellar core of radius $\sim 1$~km, so that 
the total luminosity of $\Aprime$ is 
%
\begin{equation}
L_{\mathrm{A}'} = V (Q_{\mathrm{pp}} + Q_{\mathrm{pn}})
\end{equation}
%
where $V$ is the volume of the sphere. Clearly, $L_{\mathrm{A}'}$ is proportional to $\varepsilon^2$. 
Writing $L_{\mathrm{A}'} = \varepsilon^2 I_{\mathrm{A}'}(m_{\mathrm{A}'},T)$, as in Ref.~\cite{dent_etal12}, 
and $L_{\nu}$ as the neutrino luminosity, we can write the constraint as 
%
\begin{equation}
\label{eq:bound1}
\varepsilon \lesssim \sqrt{\frac{L_{\nu}}{I_{\mathrm{A}'}(m_{\mathrm{A}'}, T)}}.
\end{equation}
%
Following SOMEBODY \arz{We need a citation here.}, we derive a constraint taking 
$L_{\nu} \approx 10^{53}\, \mathrm{erg/s} \approx 4 \times 10^{37}\, \mathrm{MeV}^2$.

The constraint derived in this manner from Eq.~(\ref{eq:bound1}) sets the lower limit on the 
exclusion band shown in Figure~\ref{fig:constraint}. We will discuss Fig.~\ref{fig:constraint} 
in more detail below. If all of the $\Aprime$ produced in the core leave the supernova freely, all 
values of $\varepsilon$ higher than those given by Eq.~(\ref{eq:bound1}) would be ruled out. 
However, as we have already mentioned, $\varepsilon$ can become sufficiently large that only a 
negligible amount of energy actually exits the supernova core. For large values or $\varepsilon$, 
this can occur because of either $\Aprime$ decays or $\Aprime$ scattering. These additional considerations place 
an upper limit on the values of $\varepsilon$ for which this constraint applies, 
and we discuss these effects in the next two subsections.

@@

The constraint from the streaming limit above provides an upper bound on the allowed coupling 
from the production on $\Aprime$ bosons. In order to function as an effective cooling channel, the 
radiated dark gauge bosons must escape the supernova in sufficient numbers. 
As the coupling increases, so do processes that prevent this escape, 
and so the excluded region has an upper bound, above which the luminosity again drops below the 
neutrino luminosity. The first such limit may be found by considering decay of the dark bosons 
into Standard Model particles. Standard model particles will scatter and thermalize 
on a timescale much shorter than the timescale for the evolution of the core, so decays 
contribute nothing to the cooling.

\arz{I removed reference to $N_{\rm eff}$ because for us $N_{\rm eff}=1$ (doesn't it? the 
only kinematically-allowed decay is to electron-positron pairs, correct?). I also rephrased a 
little bit. Please review this.}
The dark boson has a typical lifetime of
\begin{equation}
\tau_{\mathrm{A}'} = \frac{3}{\varepsilon^2 \alpha m_{\mathrm{A}'}},
\end{equation}
where $\alpha$ is the fine structure constant. Therefore, in the relativistic limit, 
the typical travel distance to decay is given by 
%
\beq
\label{eq:decaylength}
l = \gamma \tau_{\mathrm{A}'} = \frac{3 E_{\mathrm{A}'}}{N_{eff} m_{\mathrm{A}'}^2 \varepsilon^2},
\eeq
and so the fraction escaping the supernova before decaying may be estimated as (e.g., Ref.~\cite{bjorken_etal09})
\beq
\label{eq:decayfactor}
\exp \left( - \frac{r_{\mathrm{decay}}}{l} \right) = 
\exp \left( - \frac{r_{\mathrm{decay}} m_{\mathrm{A}'}^2 \varepsilon^2}{3 E_{\mathrm{A}'}} \right),
\eeq
%
an approximation which is valid so long as the size of the supernova within which standard model decay 
products can interact and be thermalized is significantly larger than the region within which $\Aprime$ are produced, 
an assumption which should be satisfied comfortably. 
To account for the suppression of gauge boson luminosity due to decays we simply multiply the phase space 
integrand in Eq.~(\ref{eq:rate1}) by the exponential suppression factor, after which the calculation proceeds 
as in the previous subsection. The limit is derived in the same way, 
with the additional complication that $I_{\mathrm{A}'}$ is now a function of $\varepsilon $, in addition to $m_{\mathrm{A}'}$ and $T$. 
The constraint Eq.~(\ref{eq:bound1}), therefore, becomes a transcendental equation that must be solved numerically.

\arz{I have a few questions here. First, what did you choose for the value or $r_{\mathrm{decay}}$? This must 
be specified.

Second, this treatment is valid under the assumption of relativistic dark photons. However, it is not clear to me that this limit is 
the appropriate limit. This limit is valid when $m_{\mathrm{A}'} < T = 30 \, \mathrm{MeV}$. However, I think that 
the decays only become more important than trapping for masses $m_{\mathrm{A}'} \gtrsim 30\, \mathrm{MeV}$. 
This suggests that we should use the non-relativistic limit here. Do I misunderstand something?}

The luminosity in $\Aprime$ will be an increasing function of $\varepsilon$ until decays suppress the gauge boson luminosity, 
at which point $L_{\mathrm{A}'}$ becomes a rapidly decreasing function of $\varepsilon$. Therefore, the excluded values of 
$\varepsilon$ at a given mass will generally have a lower bound set by the calculations of the previous subsection, and an 
upper bound set by decays. An approximate treatment of the upper bound due to decays, as we present here, is 
sufficient because over almost the entire mass range of interest, larger values of $\varepsilon$ are independently 
excluded by terrestrial beam dump experiments \cite{bjorken_etal09}. Therefore, it is far more important to 
derive an accurate estimate of the {\em lower} boundary of the exclusion region (as was done in 
the previous subsection) than the upper boundary of the exclusion region.

@@

The considerations of the previous section lead to constraints on the 
viable parameters for dark electromagnetism models of SIDM.
